# Installation

In this section, the aim is to have everyone setup with R, RStudio, the tools you need to build an R package, and __git__.

## Overview

* **Duration** 15 minutes

## Questions

* How do I install R?
* How do I install RStudio
  * What about Positron?
* How do I install git?
* How do I install RTools?

## Software Setup

### R

#### Windows

[https://cloud.r-project.org/bin/windows/](https://cloud.r-project.org/bin/windows/)

#### MacOS

[https://cloud.r-project.org/bin/macosx/](https://cloud.r-project.org/bin/macosx/)

#### Linux

[https://cloud.r-project.org/bin/linux/](https://cloud.r-project.org/bin/linux/)

### RStudio

[https://posit.co/download/rstudio-desktop/#download](https://posit.co/download/rstudio-desktop/#download)

### git

[Quarto installation page](https://quarto.org/docs/get-started/)

## Checking you are up to date

To ensure you are up to date, run the following script to install the packages.

```{r}
#| label: install-rmd
#| eval: false
install.packages("devtools")
install.packages("here")
install.packages("tidyverse")
```


## Test Script

You should be able to run the following code on your machine

```{r}
#| label: library-test
#| eval: false
library(quarto)
library(knitr)
library(here)
library(tidyverse)
library(broom)
library(fs)
library(usethis)
```
